\documentclass[9pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.2cm]{geometry}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{multicol}

\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlength{\columnsep}{12pt}

% Compact sections
\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}%
  {-4pt plus -1pt minus -1pt}{2pt plus 1pt}{\normalfont\normalsize\bfseries}}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
  {-3pt plus -1pt minus -1pt}{1pt plus 1pt}{\normalfont\small\bfseries}}
\makeatother

\begin{document}

\begin{center}
{\large\bfseries PhD Application Guide --- One-Page Summary}\\[2pt]
{\small Onat Y{\i}lmaz $\cdot$ May 2026 $\cdot$ Preprint: \textit{Event-Aligned SVD for Pattern Kernel Discovery in Crypto Perpetual Futures}}
\end{center}

\vspace{-2pt}
\hrule
\vspace{4pt}

\begin{multicols}{2}

\section*{Publishing Platforms}

\textbf{SSRN} (ssrn.com) --- No affiliation required. Free. Upload PDF, get permanent link in 1--3 days. Standard venue for finance preprints. Use \textit{Capital Markets: Market Microstructure} e-journal.

\textbf{arXiv} (arxiv.org) --- No affiliation required but needs endorsement from existing author (spam filter, not peer review). Categories: \texttt{q-fin.TR}, \texttt{q-fin.ST}, \texttt{stat.ML}.

\textbf{Strategy:} SSRN first (no barriers) $\to$ email professors with link $\to$ get arXiv endorsement $\to$ cross-post.

\section*{ETH Z\"urich}

\textbf{Josef Teichmann} -- D-MATH. ML in finance, deep hedging, rough analysis. \textit{Best overall fit.} Pitch: latent structure discovery in LOB dynamics via SVD.

\textbf{Beatrice Acciaio} -- D-MATH, SFI. Optimal transport, robust finance. Pitch: Wasserstein regime detection on order book distributions.

\textbf{Patrick Cheridito} -- D-MATH. Deep learning for optimal stopping, approximation theory. Pitch: approximation bounds on SVD-extracted features.

\textbf{Peter B\"uhlmann} -- D-MATH Statistics. High-dim inference, stability selection, FDR. Pitch: replicability guarantees for microstructure signals.

\textbf{Helmut B\"olcskei} -- D-ITET. Signal processing, harmonic analysis, ML theory. \textit{Best methods fit.} Pitch: info-theoretic limits of pattern recovery from noisy LOB data.

\textbf{Walter Farkas} -- D-MATH + UZH. Director MSc Quant Finance. Crypto theses supervised. Pitch: empirical microstructure of crypto perpetuals.

\textbf{Markus Leippold} -- UZH/SFI. NLP in finance, crypto markets. Most direct crypto expertise in Zurich. Co-supervision with ETH feasible.

\section*{EPFL Lausanne}

PhD via EDFI program (SFI). Deadlines: \textbf{Jan 15} and \textbf{Mar 31}.

\textbf{Pierre Collin-Dufresne} -- SFI. Microstructure, informed trading, order flow. \textit{Best microstructure fit.} Ex-Goldman. Pitch: SVD detection of informed trading patterns.

\textbf{Julien Hugonnier} -- SFI. Derivatives, OTC microstructure. Published ``Perpetual Futures Pricing'' (Math Finance 2025). Pitch: empirical microstructure grounded in perp pricing theory.

\textbf{Semyon Malamud} -- SFI. ML + asset pricing, ``AI Pricing Theory'' (w/ Bryan Kelly). Pitch: SVD factors as inputs to learned pricing kernels.

\textbf{Damir Filipovic} -- Swissquote Chair. Stochastic processes, ML in quant finance. Ex-Princeton, PhD ETH. Pitch: regime-switching models with SVD state variables.

\textbf{Lenka Zdeborova} -- SPOC Lab. Statistical physics of ML, phase transitions, signal recovery. Pitch: phase transitions in microstructure signal recovery. Needs SFI co-advisor.

\textbf{Florent Krzakala} -- IdePHICS Lab. Random matrix theory, spiked models, compressed sensing. Pitch: spiked matrix models for LOB feature matrices. Needs SFI co-advisor.

\textbf{Daniel Kuhn} -- RAO Lab. Distributionally robust optimization. 2024 Farkas Prize. Pitch: robust execution/sizing for pattern-based strategies.

\section*{Top Pairings}

\begin{tabular}{@{}ll@{}}
Finance + Methods: & Teichmann + B\"olcskei (ETH) \\
Microstructure + RMT: & Collin-Dufresne + Krzakala (EPFL) \\
ML pricing + Physics: & Malamud + Zdeborova (EPFL) \\
\end{tabular}

\section*{Email Template (Key Lines)}

\textit{Subject:} Prospective PhD --- Microstructure Pattern Discovery via SVD

Mention: (1) embedded systems + quant trading background, (2) preprint on SSRN with link, (3) key result: 6 significant kernels from 158 candidates after BH-FDR, turtle-soup/trap IC $\in [0.20, 0.36]$, (4) bridge to professor's specific work, (5) offer to share codebase.

\end{multicols}

\end{document}
